---
title: "House price prediction using regression analysis"
author: "RAJ KHANDAGALE ( 215280002)"
date: "15/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# House price prediction using regression analysis 



## All About Dataset 
Name of Dataset :- Real Estate data 
Source :- https://www.kaggle.com/dcw8161/real-estate-price-prediction/data
Variables :- 1) X1 transaction date (Date at which home is bought)
             2) X2 house age (age of house from when it was built)
             3) X3 distance to the nearest MRT station
             4) X4 number of convenient stores
             5) X5 latitude ( represents the geographical position of property)
             6) X6 longitude ( represents geographical position of property )
             7) Y house price of unit area
      
Dimensions :- 414 × 8       

#Problem statement :



```{r}
library(tidyverse)
```


```{r}
setwd("C:/Users/HP/Desktop/Real estate data")
```

### Reading the data as Real_df


```{r}
Real_df=read.csv('Real estate.csv' , header=TRUE)
head(Real_df)
```

```{r}
str(Real_df)
```

```{r}
dim(Real_df)
```

```{r}
summary(Real_df)
```

```{r}
sum(is.na(Real_df))
```
our data doesn't contain any N.A. values hence it is good to define a linear regression model

# model fitting 

##### baseline model 
```{r}
model1=lm(Y.house.price.of.unit.area~. , Real_df)
summary(model1)
```
considering a level of significance to be 1% 
It is found that No is just a observation number and also found to be insignificant for prediction of house price of unit area
so we define new model after removing ${\bf No. }$

###### Asumptions testing
```{r}
library(ggplot2)
library(ggfortify)
autoplot(model1)
```
### testing normality 
```{r}
shapiro.test(Real_df$Y.house.price.of.unit.area)
```
our response is not normally distributed 


```{r}
y=Real_df$Y.house.price.of.unit.area
z=sqrt(y)
shapiro.test(z)
```
at 1% l.o.s. our z=sqrt(y) satisfies normality so we will take z as response variable

#### testing homscedasticity 

```{r}
library(lmtest)
bptest(model1)
```
since p value is larger than tha 0.05 so we fail to reject the null hypothesis that data is homoscedastic. So data is homoscedastic. 

#### testing  multicolinearity

```{r}
library(mctest)
mctest(model1)
```
by determinant test we can say that multicolinearity is not detected in the model 
still for further diagnosis we can check VIF 

```{r}
library(carData)
library(car)
vif(model1)
```
since VIF for all the predictors is less than 5 so we can conclude that multicolinearity is not present in our current model 




### Model 2 
```{r}
y=Real_df$Y.house.price.of.unit.area
x1=Real_df$X1.transaction.date
x2=Real_df$X2.house.age
x3=Real_df$X3.distance.to.the.nearest.MRT.station
x4=Real_df$X4.number.of.convenience.stores
x5=Real_df$X5.latitude
x6=Real_df$X6.longitude
z=sqrt(y)
df=data.frame(z,x1,x2,x3,x4,x5,x6)
head(df)
```


```{r}
model2=lm(z~. , df)
summary(model2)
```
##### latitude(x6) found to be insignificant using p value criteria and threshold to be 5% but longitude (x5) was found significant but both together show location and single variable doesnt have a meaning so we will discard both the variables.


```{r}
model3=lm(z~.-x5-x6,df)
summary(model3)
```

```{r}
library(ggplot2)
library(ggfortify)
autoplot(model3)
```


#### Interaction terms 

```{r}
library(corrplot)
corrplot(cor(df),type="upper",method="circle",title="Correlation plot between variables",
         mar=c(0.7,0.7,0.7,0.7),tl.cex = 0.6)
```


```{r}
print(paste("from corrplot we can see that among the variables present in the model x3(distance to nearest MRT station) & x4(no of stores near to the house) has correlation between" , round(cor(x3,x4),2)))
```

since the correlation between x3 and x4 is high we will add the interaction term of x3 and x4 in the model
and also we will add higher powers of x2 and x3 so as to obtain more efficient model
```{r}
model4=lm(z~.-x5-x6+x3:x4+I(x3^2)+I(x2^3),df)
summary(model4)
```
```{r}
autoplot(model4)
```
so our current model has 67% accuracy with interaction terms of predictors and higher powers of predictor


```{r}
library(olsrr)
ols_plot_cooksd_chart(model4)
```

```{r}
ols_plot_dffits(model4)
```
20 observations are found to be outlier using Cook's D

```{r}
df1=df[-c(48,56,114,117,127,129,149,193,221,229,234,252,271,274,313,331,345,348,362,390),]
```

```{r}
df1
```
```{r}
model5=lm(z~.-x5-x6+x3:x4+I(x3^2)+I(x2^3),df1)
summary(model5)
```
```{r}
autoplot(model5)
```
since interaction term found to be insignificant in model 5 after removing outliers from the model so we will remove that and define final model

```{r}
final=lm(z~.-x5-x6+I(x3^2)+I(x2^3),df1)
summary(final)
```



since value of ${\bf R-Squared }$ is 78.24 % we conclude that Final is our  Final model


```{r}
library(formattable)
formattable(final$coefficients)
```

## Final model is 
$$
z=-825.8+0.4138\times\text{X}_{1}-0.04552\times\text{X}_{2}-0.00113\times\text{X}_{3}+0.0585\times\text{X}_{4}+1.48e-05\times \text I(\text{X}_{2}^3)+ 1.337e-07\times \text I(\text{X}_{3}^2)
$$
## Conclusion of the final model: 

1)Detected outliers in the model using Cook's D and removed them from dataset
2) From the residual vs fitted graph we can see that the estimated error curve of our final model is almost converge to 0.
3) From the QQ-Plot we can see that the our model behaves like normal except for the tail parts.
4) homoscedasticity is satisfied by the variance of residuals
5) Multi-co-linearity is not present in the final model 
6) residuals follow normal distribution with constant variance 
7) From value of $$ \text{R}^2 $$ we can conclude that final model predicts the house price per unit area ($) with 78.24% accuracy
8)Only following predictors are responsible for predicting house price 
             i) X1 transaction date (Date at which home is bought)
            ii) X2 house age (age of house from when it was built)
           iii) X3 distance to the nearest MRT station
            iv) X4 number of convenient stores from house


